### 2025-04-17 closed_loop_planning.tex
[MGa] textual description of closed-loop planning

\section{Closed-loop uncertainty-aware planning over the full horizon}
\label{sec:closed_loop_planning}

The second approach plans over the full horizon while incorporating a feedback policy that mimics the driver's closed-loop behavior. This enables a more realistic propagation of the uncertainty encoded in the covariance matrix $\bP$, preventing its uncontrolled growth through the stabilizing effect of the driver's feedback.

The approach is based on the three following steps: i) A nominal time-optimal trajectory is planned: this will be referred to as nominal feed-forward optimal trajectory. This is the trajectory a driver should follow in the ideal situation of a world without uncertainty. Therefore, in this step the mean trajectory is the actual trajectory a deterministic vehicle dynamics would trace. ii) A closed-loop controller mimicking the driver's action is computed which stabilizes the nominal feed-forward optimal trajectory computed at step i). We assume a discrete time-varying Linear Quadratic Regulator (LQR) controller is a good approximation of how an expert driver can track a prescribed trajectory. iii) In the last step the stochastic framework is reintroduced and the time-optimal planning problem is robustified using the Lyapunov framework. The main idea is to \emph{re-plan} a mean trajectory of a stochastic system such that, a time-varying LQR controller that aims at stabilizing it, is able to properly tame the propagation of the uncertainty. This amounts to ensuring that the robustified path constraints are satisfied. More specifically, this implies two key aspects: first, that the covariance matrix is propagated accounting explicitly for the controlled system dynamics; and second, that the path constraints are robustly satisfied by incorporating the covariance estimation propagated through the closed-loop controlled system. Some aspects that may initially appear technical are in fact essential. In this third step, not only are states and controls re-planned, but also the time-varying LQR gains. In fact, they need to explicitly account for the changed requirements associated with stabilizing a trajectory different from the nominal one, which must now satisfy the robustified constraints.
\subsection{Suspension Analysis}
\label{sec:suspanal}
The suspensions connect the chassis to the unsprung masses. This connection cannot be described by an elementary joint, such as prismatic or revolute. Therefore, a set of six variables $q_{bh}$, that parameterize a homogeneous transformation matrix, is necessary. The transformation can be realized by composing three translations and three rotations (the ZXY Euler sequence was selected) using the global Product of Exponentials (PoE) formula~\cite{Murray:book:1994}, which reads: \begin{equation}\label{eqn:PoE}
	g_{bh}=e^{\hat{Y}_1q_{bh,1}}\cdots e^{\hat{Y}_6q_{bh,6}} g_{bh}(0),
\end{equation}
where the offset $g_{bh(0)}=I$ and $Y_i=e_i$, with $e_i\in\bbR^6$ are the six normalized screw vectors (canonical basis elements in $\bbR^6$) associated with the Cartesian coordinates and the Euler ZXY angles in $q_{bh}$.
Clearly, the six components of $q_{bh}$ are not independent; the modeled suspension is, in fact, a double wishbone that has: 2 DoFs for wheels that can steer, and 1 DoF otherwise.
In this case, both axles have wheels that can steer; hence, a pair of independent variables ($z;\delta $) is associated with each wheel.
A detailed analysis is omitted here for the sake of brevity. The interested reader is referred to~\cite{Domenighini:Designs:2023} for a similar treatment using the PoE formalism.

%As mentioned in the previous sections, $z$ represents the variation of the distance between the two attachment points of the spring-damper system, positive when the distance increases and negative when it decreases, while $\delta$ represents the motion of the steering rod attachment point to the chassis along its local $y$ axis (frame $\sref{B}$). Considering that the type of the suspension is known, six scalar equations of constraints can be expressed implicitly as $F(q_{bh};z,\delta)=0\in \bbR^6$.
%
%To find an explicit relation between inputs $(z,\delta)$ and outputs $q_{bh}$, the constraint equations are solved on a grid of pairs $(z^{(p)},\delta^{(l)})$, where $p = 1,\dots,N$ and $l = 1,\dots,M$. Then, the data are fitted (in the least-squares sense) with a 2D regression polynomial $q_{bh}(z,\delta)$ as follows:
%\begin{equation}\label{eqn:poly}
%	q_{bh,i}(z,\delta)=\sum_{r=0}^{3} \sum_{s=0}^{3-r} a_{{rs},i} z^r \delta^s.
%\end{equation}
%For the purpose a polynomial of degree three is sufficiently accurate. Now is possible to express the homogeneous transformation $g_{bh}\bigl(q_{bh}(z,\delta)\bigr)$ and to compute the twist $V_{bh}$ and $\dot{V}_{bh}$ as functions of $(z,\delta)$ and their derivatives. We start from:
%\begin{equation}\label{eqn:Jbh}
%	V_{bh}=\sum_{i=1}^6J_{bh,i}\dot{q}_{bh,i},
%\end{equation}
%where $J_{bh}$ is the \emph{geometric Jacobian} of the suspension joint relative to $q_{bh}$. Then, following the procedure thoroughly described in~\cite[Sec. 2.2]{Domenighini:Designs:2023}, it is possible to obtain %
%assumed with independent components. Each column of $J_{bh}$ can be computed based on the PoE formula \eqref{eqn:PoE}. For the $i$-th column, we have $J_{bh,i}=A_i^{-1}Y_i$, with $A_6 = I_{6\times 6}$ and $A_i=\Ad_{e^{\hat{Y}_{i}q_{bh,i}}\cdots e^{\hat{Y}_6q_{bh,6}}}$, $i = 1, \dots, 5$. The \emph{Adjoint operator} $\Ad_g$ has been defined as the $6\times6$ matrix satisfying $Y=\Ad_gX$ whenever $\hat{Y}=g\hat{X}g^{-1}$, so that, if $g=\left[\begin{smallmatrix}R&d\\0^T&1\end{smallmatrix}\right]$, then $\Ad_g=\left[\begin{smallmatrix}R&\hat{d}R\\0&R\end{smallmatrix}\right]$.\\
%As a function of $(z,\delta)$, $\dot{q}_{bh,i}$ can be expressed as follows:
%\begin{equation}\label{eqn:dt_qbh}
%	\dot{q}_{bh,i}=\frac{\partial}{\partial z}q_{bh,i}\dot{z}+\frac{\partial}{\partial\delta}q_{bh,i}\dot{\delta},
%\end{equation}
%where the (geometric) partial derivatives $\frac{\partial}{\partial z}q_{bh,i}$ and $\frac{\partial}{\partial\delta}q_{bh,i}$ are obtained by differentiating Eq.~\eqref{eqn:poly}.
%Then, by defining the Jacobians with respect to the true independent variables $z$ and $\delta$ as follows
%\begin{align}\label{eqn:Jbh,z}
%	J_{bh,z}=\sum_{i=1}^6J_{bh,i}\frac{\partial}{\partial z}q_{bh,i}\quad \textrm{and}\quad J_{bh,\delta}=\sum_{i=1}^6J_{bh,i}\frac{\partial}{\partial\delta}q_{bh,i},
%\end{align}
%it possible to obtain the rigid-body velocity $V_{bh}$ of the constrained motion from Eq.~\eqref{eqn:Jbh}, in terms of $\dot{z}$ and $\dot{\delta}$, as follows
%\begin{equation}\label{eqn:Vbh}
%	V_{bh}=J_{bh,z}\dot{z}+J_{bh,\delta}\dot{\delta}.
%\end{equation}
%
%To write the dynamic equations the time derivative $\dot{V}_{bh}$ of the twist $V_{bh}$ is necessary. To this sake, it is sufficient to differentiate Eqs.~\eqref{eqn:Jbh}--\eqref{eqn:Vbh}.
%\begin{equation}\label{eqn:dt_Jbh}
%	\dot{V}_{bh}=\sum_{i=1}^6\dot{J}_{bh,i}\dot{q}_{bh,i}+\sum_{i=1}^6J_{bh,i}\ddot{q}_{bh,i}.
%\end{equation}
%
%Here, following~\cite{Gabiccini:ICRA:2012}, the derivatives $\dot{J}_{bh,i}$ can computed using the formula
%\begin{align}\label{eqn:dt_Jbhi}
%	\dot{J}_{bh,i}=-\sum\limits_{i < j \leq 6}\ad_{J_{bh,j}\dot{q}_{bh,j}} J_{bh,i}
%\end{align}
%using Lie derivatives. In the matrix case, the Lie derivative $\hat{Z}$ of the vector field $\hat{X}$ with respect to vector field $\hat{Y}$ is calculated by as a commutator operation $\hat{Z}=\hat{Y}\hat{X}-\hat{X}\hat{Y}$; in the vector case, the $6\times6$ matrix $\ad_Y$ is used, defined such that $Z=\ad_Y X$. If $v$ and $\omega$ are the translational and rotational component of $V$, then $\ad_V=\left[\begin{smallmatrix}\hat{\omega}&\hat{v}\\0&\hat{\omega}\end{smallmatrix}\right]$.
%
%In Eq.~\eqref{eqn:dt_Jbh}, the accelerations $\ddot{q}_{bh,i}$ are obtained by differentiating the velocities in Eq.~\eqref{eqn:dt_qbh}:
%\begin{equation}\label{eqn:dt_dt_qbh}
%	\ddot{q}_{bh,i}=\frac{\partial^2}{\partial z^2}q_{bh,i}\dot{z}^2+2\frac{\partial^2}{\partial z\partial\delta}q_{bh,i}\dot{z}\dot{\delta}+\frac{\partial^2}{\partial\delta^2}q_{bh,i}\dot{\delta}^2+\frac{\partial}{\partial z}q_{bh,i}\ddot{z}+\frac{\partial}{\partial\delta}q_{bh,i}\ddot{\delta},
%\end{equation}
%where the partial derivatives are computed by differentiating the regression polynomials~in~\eqref{eqn:poly}.
%After simple but tedious calculations, the explicit expression for the derivatives of the Jacobians in \eqref{eqn:Jbh,z} appear as follows:
%\begin{align}
%	\dot{J}_{bh,z}&=\sum_{i=1}^6\dot{J}_{bh,i}\frac{\partial}{\partial z}q_{bh,i}+\sum_{i=1}^6J_{bh,i}\Bigl(\frac{\partial^2}{\partial z^2}q_{bh,i}\dot{z}+\frac{\partial^2}{\partial z\partial\delta}q_{bh,i}\dot{\delta}\Bigr)\label{eqn:dt_Jbh,z}\quad\textrm{and}\\\dot{J}_{bh,\delta}&=\sum_{i=1}^6\dot{J}_{bh,i}\frac{\partial}{\partial\delta}q_{bh,i}+\sum_{i=1}^6J_{bh,i}\Bigl(\frac{\partial^2}{\partial z\partial\delta}q_{bh,i}\dot{\delta}+\frac{\partial^2}{\partial\delta^2}q_{bh,i}\dot{\delta}\Bigr)\label{eqn:dt_Jbh,d}.
%\end{align}
%Substituting Eqs.~\eqref{eqn:dt_Jbhi} and~\eqref{eqn:dt_dt_qbh} in Eqs.~\eqref{eqn:dt_Jbh}, and casting some intermediate results as~\eqref{eqn:dt_Jbh,z} and~\eqref{eqn:dt_Jbh,d},
%the constrained rigid-body acceleration $\dot{V}_{bh}$ in the following form:
%\begin{equation}\label{eqn:dt_Vbh}
%	\dot{V}_{bh}=\dot{J}_{bh,z}\dot{z}+\dot{J}_{bh,\delta}\dot{\delta}+J_{bh,z}\ddot{z}+J_{bh,\delta}\ddot{\delta}.
%\end{equation}
%The explicit expressions for auxiliary Jacobians $J_{bh,z}$, $J_{bh,\delta}$, $\dot{J}_{bh,z}$ and $\dot{J}_{bh,\delta}$ are omitted here for brevity. The interested reader can find the detailed analytical derivation in~\cite[Eqs. (5)--(8)]{Domenighini:Designs:2023}.
%
%A generalized force $\tau\in\bbR$ is introduced that depends only on the independent variable $z\in\bbR$. $\tau$ is assumed to be generated only by the spring and the damper, considered aligned along the same axis. It is worth noting that, using the defined $z$, we are allowed to compute the force of the suspension without calculate the associated $q_{bh}$ nor using the Principle of Virtual Works. In this expression, also the contribution of the anti-roll bar is considered, which depends on the difference between the left and right trim, and the contribution of the bump-stops, which activate when the trim reaches its lower or upper limits. The intensity of $\tau$, for each wheel, is expressed as follow:
%\begin{equation}
%	\tau_{ij}=k_{s,i}(l_{c,i}-l_{0,i}+z_{ij})+c_{d,i}\dot{z}_{ij}+(-1)^j k_{a,i}(z_{i,1}-z_{i,2}) + \left(\frac{2z-z_{\text{max}}-z_{\text{min}}}{|z_{\text{max}}-z_{\text{min}}|}\right)^{b},
%\end{equation}
%where $l_{c,i}$ is the spring's length in the reference position, $l_{0,i}$ is the spring's length at rest, $k_{s,i}$ and $c_{d,i}$ are the stiffness and the damping coefficients, $k_{a,i}$ is the anti-roll bar stiffness, $b$ influences the stiffness of the bump-stops, and $z_{\text{min}}$ and $z_{\text{max}}$ are the knees of the bump-stops' force characteristics.
\section{Tire Model}
\label{sec:tyremodel}
\begin{figure}[t]\centering
	\tdplotsetmaincoords{0}{0}
	\begin{tikzpicture}[tdplot_main_coords,scale=3.5]\small
		\tdplotsetrotatedcoords{-166.5322}{-65.2292}{-166.5322}\draw[tdplot_rotated_coords,dashed,gray!75,fill=gray!10](-0.6,0,0)--(-0.6,1.1,0)--(0.6,1.1,0)--(0.6,0,0);\draw[tdplot_rotated_coords,very thick,fill=white](0,0)coordinate(C)..controls(0.25,0)and(0.5,0)..(0.5,0.45)arc(0:180:0.5)..controls(-0.5,0)and(-0.25,0)..(0,0);\draw[tdplot_rotated_coords,very thick,fill=gray!25](0,0.45)coordinate(O)circle(0.25);\draw[tdplot_rotated_coords,very thick,dashed](0.5,0.45)arc(0:-180:0.5);\draw[tdplot_rotated_coords,very thin,dashed](0,0,0)--(0,0.2,0)(0,0.7,0)--(0,0.95,0)(-0.5,0.45,0)--(-0.25,0.45,0)(0.25,0.45,0)--(0.5,0.45,0);\draw[tdplot_rotated_coords,very thin](0,0.2,0)--(0,0.7,0)(0,0.95,0)--(0,1.25,0)(-1,0.45,0)--(-0.5,0.45,0)(-0.25,0.45,0)--(0.25,0.45,0)(0.5,0.45,0)--(1,0.45,0);\draw[tdplot_rotated_coords,{Stealth[inset=0pt]}-{Stealth[inset=0pt]},very thin](0.75,0,0)--node[right]{$r$}(0.75,0.45,0);\tdplotsetrotatedcoordsorigin{(O)}\tdplotsetrotatedcoords{-166.5322}{-65.2292}{-146.5322}\draw[tdplot_rotated_coords,dashed](-0.25,0,0)--(-0.5,0,0);\draw[tdplot_rotated_coords,dashed](0,0.25,0)--(0,0.5,0);\draw[tdplot_rotated_coords,->](0,0,0)--(-0.25,0,0)(-0.5,0,0)--(-1,0,0)node[below left]{$x_h$};\draw[tdplot_rotated_coords,->](0,0,0)--(0,0.25,0)(0,0.5,0)--(0,0.8,0)node[above left]{$z_h$};\draw[tdplot_rotated_coords,->](0,0,0)--(0,0,0.8)node[right]{$y_h$};\draw[tdplot_rotated_coords,{Stealth[inset=0pt]}-{Stealth[inset=0pt]},very thin](-0.8,0)arc(180:160:0.8)node[midway,left]{$\sigma$};\tdplotsetrotatedcoordsorigin{(C)}\tdplotsetrotatedcoords{43.2192}{-27.9909}{-46.7808}\draw[tdplot_rotated_coords,dashed,gray!75,fill=gray!25](0,0,-0.6)--(0.6,0,-0.6)--(0.6,0,0.6)--(0,0,0.6);\draw[tdplot_rotated_coords,->](0,0,0)--(0.8,0,0)node[right]{$y_n$};\draw[tdplot_rotated_coords,->](0,0,0)node[circle,fill=black,inner sep=0pt,minimum size=3pt,draw]{}--(0,1.25,0)node[above]{$z_n$};\draw[tdplot_rotated_coords,->](0,0,-1)--(0,0,1)node[below left]{$x_n$};\draw[tdplot_rotated_coords,{Stealth[inset=0pt]}-{Stealth[inset=0pt]},very thin](0,1.1)arc(90:110:1.1)node[midway,above]{$\gamma$};\draw(0,-0.02)..controls(0,-0.3)and(0.3,-0.1)..(0.3,-0.3)node[below]{contact point};\draw(-0.5,0.7)..controls(-0.35,0.55)and(-0.6,0.7)..(-0.55,0.5)node[below,align=center]{wheel\\plane};\draw(0.4,0.15)..controls(0.3,0.25)and(0.55,0.2)..(0.5,0.3)node[above,align=center]{road\\plane};\draw(O)node[below right]{$O_h$}(C)node[below right]{$O_n$};\draw[tdplot_rotated_coords,thick,-{Stealth[inset=0pt]}](C)--+(0,0,0.8)node[below right]{$F_x$};\draw[tdplot_rotated_coords,thick,-{Stealth[inset=0pt]}](C)--+(0.5,0,0)node[above]{$F_y$};\draw[tdplot_rotated_coords,thick,-{Stealth[inset=0pt]}](C)--+(0,1,0)node[right]{$F_z$};
	\end{tikzpicture}
	\hfill
	\begin{tikzpicture}[scale=4.25]\small
		\coordinate(C)at(0,0);\coordinate(-)at(-0.375,0);\coordinate(+)at(0.325,0);\draw[->](-0.7,0)--(0.6,0)node[right]{$y_n$};\draw[->](0,0)node[circle,fill=black,inner sep=0pt,minimum size=3pt,draw]{}--(0,1.25)node[above]{$z_n$};\tdplotsetrotatedcoords{10}{0}{0}\draw[tdplot_rotated_coords,dash dot](0,-0.15)--(0,1.25)(-0.35,0.45)--(0.6,0.45);\draw[tdplot_rotated_coords,very thick](-0.25,0.2)--(-0.25,0.7)..controls(-0.25,0.95)and(-0.25,0.95)..(0,0.95)..controls(0.25,0.95)and(0.25,0.95)..(0.25,0.7)--(0.25,0.2)(-0.25,0.2)..controls(-0.25,0.025)and(-)..(C)..controls(+)and(0.25,-0.025)..(0.25,0.2);\draw[tdplot_rotated_coords,very thick,dashed](-0.25,0.2)..controls(-0.25,-0.05)and(-0.25,-0.05)..(0,-0.05)node(B){}..controls(0.25,-0.05)and(0.25,-0.05)..(0.25,0.2);\node[tdplot_rotated_coords](O)at(0,0.45){};\draw[very thin](O)--({(-0.55,0)}|-O);\draw[very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](-0.475,0)--node[left]{$h$}({(-0.475,0)}|-O);\draw[very thin](B)--({(-0.7,0)}|-B);\draw[very thin,-{Stealth[inset=0pt]}](-0.625,0.3)--node[left]{$d$}(-0.625,0);\draw[very thin](-0.625,0)--({(-0.625,0)}|-B);\draw[very thin,-{Stealth[inset=0pt]}](-0.625,-0.15)--({(-0.625,0)}|-B);\draw[tdplot_rotated_coords,very thin](0.05,0)--(0.45,0);\draw[tdplot_rotated_coords,very thin](0.05,-0.05)--(0.6,-0.05);\draw[tdplot_rotated_coords,very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](0.375,0)--node[right]{$r$}(0.375,0.45);\draw[tdplot_rotated_coords,very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](0.525,-0.05)--node[right]{$r_0$}(0.525,0.45);\draw[very thin,{Stealth[inset=0pt]}-{Stealth[inset=0pt]}](0,1.125)arc(90:100:1.125)node[midway,above]{$\gamma$};\draw(0,-0.4);
	\end{tikzpicture}
	\caption{Decomposition of the tyre forces along the axes of the auxiliary frame $\{N\}$. The frame $\{N\}$ is such that: $x_n$ lies along the intersection between the wheel plane ($x_hz_h$-plane) and the road plane ($x_s y_s$-plane), pointing forward; $y_n$ points left (w.r.t. the driver) along the intersection line between the road plane ($x_s y_s$-plane) and the transverse plane ($y_h z_s$-plane) through the wheel center $O_h$; $z_n$ is normal to the road surface. The contact point is estimated in correspondence to the origin $O_n$. On the right, we report a detail of the tyre vertical deformation.}
	\label{fig:tyre_wheel}
\end{figure}

The tangential forces exchanged between the tire and road are computed using \mbox{Pacejka's} \emph{Magic Formula}:
\begin{equation}\label{eqn_mf}
	[F_x,F_y]=\mf(\kappa,\alpha,F_z,\gamma),
\end{equation}
where $ F_x, F_y $ are the longitudinal and lateral forces, dependent on tire slips $( \kappa, \alpha )$, vertical load $( F_z )$, and camber angle $( \gamma )$.
As shown in Fig.~\ref{fig:tyre_wheel}, an auxiliary frame $\sref{N}$ is introduced to calculate $ F_z $. This depends upon the interpenetration $ d $ of the tire into the ground, following a penalty-based compliant tire model. This model treats the wheel as a radial spring-damper system with constant stiffness $ k_t $ and damping $ c_t $. The vertical force is expressed as:
\begin{equation}\label{eqn_Fz}
	F_z= F_0\log_2\left(1 + 2^{\frac{k_t d + c_t \dot{d}}{F_0}}\right)
\end{equation}
which is differentiable for $ d = 0 $ and $\dot{d} = 0$, improving numerical stability. The model ensures $ F_z $ approaches zero when the tire loses contact with the ground, while maintaining a non-zero gradient to guide optimization algorithms in scenarios where the tire detaches from the ground or encounters sudden changes.

%To compute the tangential components of the force exchanged by tyre and road, a formulation of the Pacejka's \emph{Magic Formula} ~\cite[Section~4.3.2]{Pacejka:book:2012} is used, which reads:
%\begin{equation}\label{eqn_mf}
%	[F_x,F_y]=\mf(\kappa,\alpha,F_z,\gamma).
%\end{equation}
%The Magic Formula $\mf$ computes the longitudinal and lateral forces $F_x,F_y$ as a function of the \emph{tyre slips} $\kappa,\alpha$~\cite[Section~1.2.1]{Pacejka:book:2012}. The formula is also sensitive to variations of vertical load $F_z$ and camber angle $\gamma$.
%
%The auxiliary frame $\sref{N}$ is introduced to compute quantities used in Eq.~\eqref{eqn_mf} and the (attempted) interpenetration $d$ (see Figure.~\ref{fig:tyre_wheel}) of the tyre in the ground, necessary to compute the dynamic vertical load $F_z$. The construction of the frame $\sref{N}$ is reported in Figure~\ref{fig:tyre_wheel}. It is worth noting that the frame $\sref{N}$ is fully described if $\sref{H}$ and $\sref{S}$ are known.
%The model used for the contact is a unilateral, penalty-based compliant tyre model, that considers the wheel as a radial spring with constant stiffness $k_t$ and constant damping coefficient $c_t$\footnote{The introduction of a more refined model with nonlinear spring and damping characteristics would not pose particular difficulties.}. The traditional role of compression in a spring is conducted by $d$, which is computed considering the road and the lowermost point of the non-deformed tyre. The resultant function that relate $d$, $\dot{d}$ and $F_z$ is:
%\begin{equation}\label{eqn_Fz}
%	F_z= F_0\log_2\left(1 + 2^{\frac{k_t d + c_t \dot{d}}{F_0}}\right)
%\end{equation}
%This function, with $F_0$ tuned to be negligible, is differentiable for $d=0$ and $\dot{d}=0$, enhancing numerical performance, and behaves as a spring-damper system, as requested, when the tyre is in contact with the ground. When the tyre detaches completely from the ground, the resultant vertical forces tends to zero. However, since its gradient is not zero, this formulation helps the optimization algorithm to figure out proper search directions. This allows to deal with jumps and other situations in which the tyre may lose contact with the ground.

\subsection{Camber System Modeling}
\label{sec:cambersystemmodeling}
The dynamic camber control (DCC) system is modeled as a simplified actuated revolute joint connecting the two virtual halves of the knuckle, though a real implementation might require a more complex 1-DoF joint. While this simplification may influence joint values from the optimizer, it results in nearly the same optimal wheel posture. The choice was driven by the limited literature on such systems for FSAE vehicles, which is the vehicle type employed in our analysis. A realistic model, if available, could be characterized using the same process applied to the suspension. During the design phase of the actual joint, some parameters could be treated as time-invariant optimization variables, constrained within dimensional bounds, to achieve the best performance configuration.
%The dynamic camber control (DCC) system is modeled here as a simple actuated revolute joint connecting the two virtual halves of the knuckle. A real implementation could require a more complex generalized 1-DoF joint. Changing the joint characteristics might affect the joint values obtained from the optimizer but would result in approximately the same optimal wheel posture. This simplification was necessary due to the limited open literature on such systems for FSAE vehicles. If a realistic model of the system were available, the process used for the suspension could easily be repeated to characterize this joint. It is worth observing that, in the design phase of the actual controlled joint, some model parameters could be treated as time-invariant optimization variables, appropriately constrained to meet dimensional bounds, and optimized to find the best performance configuration.

In Table~\ref{tab:camber_prop}, we report the range of variation of function $q_{hk}(t)$ and its first and second derivatives. Considering the scarcity of open literature available, reasonable values have been selected. In Figure~\ref{fig:camber_showing} the wheel is shown in a general posture with zero camber (shaded black) and with maximum positive camber (solid black). Due to the convention adopted for rotations, this posture corresponds to the minimum value of $q_{hk}$.
\begin{table}[h]
	\centering
	\begin{tabular}{|c|c|c|}
		\hline
		$q_{hk}$ (rad) & $\dot{q}_{hk}$ (rad/s)& $\ddot{q}_{hk}$ (rad/s$^2$)\\\hline
		$\pm \pi/20$ & $\pm \pi/20$ & $\pm \pi/10$ \\\hline
	\end{tabular}
	\caption{Upper and lower bounds for camber angle value $q_{hk}(t)$ and its derivatives.}
	\label{tab:camber_prop}
\end{table}

\begin{figure}[h]
	\centering
	\includegraphics[scale = .6]{Images/camber_showing}
	\caption{Real scale representation of the wheel in two configurations: $q_{hk}=0$ (shaded surface) and $q_{hk}=-\pi/20$ (solid surface).}
	\label{fig:camber_showing}
\end{figure}
\subsection{External Wrenches}
\label{sec:extwren}
The external wrenches acting on each body, from wheels to chassis, are expressed in body-fixed components. Only gravity, aerodynamic forces, and road-tyre interactions are considered. Following~\cite{Guiggiani:book:2023}, the aerodynamic wrench in frame $\{B\}$ is $W_a=-\frac{1}{2}\rho S v_{gb,1}^2 [C_x, 0, C_z, 0, -h_0 C_x - a_f C_{zf} + a_r C_{zr}, 0]^T$, where $\rho$ is air density, $S$ is the frontal area, $v_{gb,1}$ the forward velocity (component of \emph{distal rigid-body twist} $V_{gb}$, see~\cite[p.~4]{Domenighini:Designs:2023}), and $C_x, C_z$ are drag/lift coefficients. $C_z$ is split into $C_{zf}=k_a C_z$ and $C_{zr}=(1-k_a)C_z$ by the aerodynamic balance coefficient $k_a$.

The total external wrench on a wheel is $W_{we}=\Ad_{g_{hn}}^*W_t+W_{w_h}$,
%\begin{equation}\label{eqn:Wwe}
%W_{we}=\Ad_{g_{hn}}^*W_t+W_{w_h},
%\end{equation}
where $W_t=[F_x, F_y, F_z, 0, 0, 0]^T$ is the ground wrench transferred from $\sref{N}$ to $\sref{H}$ by the \emph{starred} Adjoint operator, and $W_{w_h}=[F_{w_h}^T R_{hg}^T, 0, 0, 0]^T$ is the weight contribution with $F_{w_h}=[0, 0, -m_h g]^T$ in frame $\{G\}$, where $m_h$ is the wheel mass.

Knuckles experience no external forces, therefore:
\begin{equation}\label{eqn:Whe}
	W_{he}=[0,0,0,0,0,0]^T\quad\textrm{and}\quad W_{ke}=[0,0,0,0,0,0]^T.
\end{equation}
The chassis's external wrench is $W_{be}=W_a+W_{w_b}$, %\begin{equation}\label{eqn:Wbe}
%	W_{be}=W_a+W_{w_b},
%\end{equation}
where $W_a$ and $W_{w_b}$ are aerodynamic and weight contributions, respectively. For $W_{w_b}=[F_{w_b}^T R_{bg}^T, 0, 0, 0]^T$, the weight force is $F_{w_b}=[0, 0, -m_b g]^T$ in frame $\{G\}$, with $m_b$ as the chassis mass.
%The purpose of this section is to describe the external wrenches acting on each body, from the wheels to the chassis, each expressed in body-fixed components.
%Besides gravity, aerodynamic forces and the interactions between road and tyres are the only external forces considered.
%Adopting the notation in~\cite{Guiggiani:book:2023}, the aerodynamic contribution to the wrenches in frame $\{B\}$ is $W_a=-\frac{1}{2}\rho Sv_{gb,1}^2[C_x,0,C_z,0,-h_0C_x-a_fC_{zf}+a_rC_{zr},0]^T$, where $\rho$ is the air density, $S$ the frontal area of the vehicle, $v_{gb,1}$ the forward velocity, $h_0$ the nominal height of the CoM from ground and $a_f,a_r$ the distances of the CoM from the front and rear axle, respectively. The drag and lift coefficient $C_x$ and $C_z$ are dimensionless parameters that depend on the shape of the vehicle's body. $C_z$ can be further partitioned into $C_{zf}=k_a C_z$ and $C_{zr}=(1-k_a) C_z$ according to the \emph{aerodynamic balance coefficient} $k_a$, which is a parameter of the vehicle set-up.
%
%The total external wrench acting on a (general) wheel is:
%\begin{equation}\label{eqn:Wwe}
%	W_{we}=\Ad_{g_{hn}}^*W_t+W_{w_h},
%\end{equation}
%where $W_t=[F_x,F_y,F_z,0,0,0]^T$ are the ground forces computed in $\sref{N}$ and properly transferred to $\sref{H}$ components by the \emph{starred} Adjoint operator, while $W_{w_h} = [ F_{w_h}^T R^T_{hg}, 0,0,0 ]^T$, with $F_{w_h}=[0,0,-m_h g]^T$ in frame $\{G\}$, represents the weight contribution and $m_{h}$ is the wheel mass.
%
%Under our hypotheses, the knuckles are not subjected to external forces, thus \begin{equation}\label{eqn:Whe}
%	W_{he}=[0,0,0,0,0,0]^T\quad\textrm{and}\quad W_{ke}=[0,0,0,0,0,0]^T.
%\end{equation}
%
%The resultant external wrench acting on the chassis is therefore:
%\begin{equation}\label{eqn:Wbe}
%	W_{be}=W_a+W_{w_b},
%\end{equation}
%where $W_a$ and $W_{w_b}$ are, respectively, the contributions of aerodynamic and chassis weight forces, whose general expression is of the form $W_{w_b} = [ F_{w_b}^T R^T_{bg}, 0,0,0 ]^T$. Explicitly we have $F_{w_b}=[0,0,-m_b g]^T$ in frame $\{G\}$, where $m_b$ is the mass of the chassis.

\subsection{Recursive computation of the dynamics equations with ABA}
\label{sec:aba}
Thanks to the ABA algorithm, assuming known joint positions and velocities and joint forces and torques, it is possible to compute recursively and efficiently the accelerations of the joint variables. Following~\cite{Domenighini:Designs:2023}, twists are streamlined as follows: $V_{gb}^b = V_{gb} = V_b$, $V_{bh}^h=V_{bh}$, $V_{hw}^h=V_{hw}$. For the second model (DCC) the last expression is replaced with $V_{kw}^k=V_{kw}$ and $V_{hk}^k=V_{hk}$ is added. Given that the first model (without camber control) is a particular case of the DCC one, to avoid duplications only the algorithm for DCC is reported here.

The traditional ABA requires knowledge of the active forces and torques, denoted as $\tau_{ij}$, exerted by the parent body ($i$) on the child body ($j$). Its purpose is to solve the \emph{Forward Dynamics} problem of an articulated systems of rigid bodies. In our floating-base system, given the twist $V_b$ of the parent body $\{B\}$, the values and time derivatives of all joint variables ($q$ and $\dot{q}$), and the torque vector ($\tau$), the objective is to compute the joint accelerations ($\dot{V}_{b}$ and $\ddot{q}$).
%The fundamental concept is that of the inertia of an articulated body. This is represented by that of the root rigid body (original parent) plus the portion of the sub-trees, composed by the children bodies, whose inertia that can be structurally transferred backwards through the joints.
%This depends on the architecture, the pose and the joints' types.
The overall computation is performed in three sweeps defined recursively. One of the greatest computational benefits when it is employed to assemble symbolic dynamic equations, as in our case, is that \emph{no matrix inversion is required}, thus contributing to streamline the algebraic expressions.

The first step of the algorithm, as described in \textbf{Step 1}, performs the forward propagation of rigid-body velocities: from the root to the leaves of the tree.
%Twists are calculated as the sum of parent twist and the relative twist which is expressed consistently with the joint type. The interesting Jacobians are listed here: $J_{hk}=[0,0,0,1,0,0]^T$ and $J_{kw}=[0,0,0,0,1,0]^T$, and represent two revolute joints with axes $x$ and $y$ respectively.
In Algorithm~\ref{alg:step1} the terms ``Knuckle 1'' and ``Knuckle 2'' are used to identify the two halves of the knuckle, where, using kinematic trees terminology, the first is the parent of the second.
The Adjoint operator is defined as follows:
\begin{equation}\label{eq:Adjoint} \Ad_{g}=\left[\begin{array}{cc}R&\hat{d}R\\0&R\end{array}\right].
\end{equation}
In the case $g_{bh}\in SE(3)$ is considered, $R_{bh}\in SO(3)$ and $d_{bh}\in \bbR^3$ are functions of $(z,\delta)$. In the case of $g_{hk}$, the expressions are simply $d_{hk}=[0,0,0]^T$ and $R = \rotX(q_{hk})$, the latter being an elementary rotation about the common $x$-axis of the angle $q_{hk}$.

\begin{algorithm}[h]
	\floatname{algorithm}{Step}
	\caption{Forward Propagation of Velocity}\label{alg:step1}
	\begin{algorithmic}[1]
		\vspace{1mm}
		\For{$h=h_1,h_2,h_3,h_4$}
		\vspace{1mm}
		\State{$V_{bh}=J_{bh,z}\dot{z}+J_{bh,\delta}\dot{\delta}$}
		\vspace{1mm}
		\State{$V_h=\Ad_{g_{bh}}^{-1}V_b+V_{bh}$}\Comment{Knuckle 1 Rigid-Body Velocity}
		\vspace{1mm}
		\State{$V_{hk} = J_{hk}\dot{q}_{hk}$}
		\vspace{1mm}
		\State{$V_k=\Ad_{g_{hk}}^{-1}V_h+V_{hk}$}\Comment{Knuckle 2 Rigid-Body Velocity}
		\vspace{1mm}
		\State{$V_{kw}=J_{kw}\omega$}
		\vspace{1mm}
		\State{$V_w=V_k+V_{kw}$}\Comment{Rim Rigid-Body Velocity}
		\vspace{1mm}
		\EndFor
		\vspace{1mm}
	\end{algorithmic}
\end{algorithm}

The second step of the algorithm, as detailed in \textbf{Step 2}, performs the backward propagation of inertia and bias terms.
\begin{algorithm}[h]
	\floatname{algorithm}{Step}
	\caption{Backward Propagation of Articulated Inertia and Bias}\label{alg:step2}
	\begin{algorithmic}[1]
		\vspace{1mm}
		\For{$h=h_1,h_2,h_3,h_4$}
		\vspace{1mm}
		\State{$\hat{M}_w=M_w$}\Comment{Rim Articulated Inertia}
		\vspace{1mm}
		\State{$\hat{B}_w=-W_{we}+\ad_{V_w}^*M_wV_w$}\Comment{Rim Articulated Bias}
		\vspace{1mm}
		\State{$\bar{M}_w=\hat{M}_w-\dfrac{\hat{M}_wJ_{hw}J_{hw}^T\hat{M}_w}{J_{hw}^T\hat{M}_wJ_{hw}}$}
		\vspace{1mm}
		\State{$\bar{B}_w=\hat{B}_w-\dfrac{\hat{M}_wJ_{hw}J_{hw}^T\hat{B}_w}{J_{hw}^T\hat{M}_wJ_{hw}}-\bar{M}_w\ad_{V_{hw}}V_w$}
		\vspace{1mm}
		\State{$\hat{M}_k= \bar{M}_w$
		}\Comment{Knuckle 2 Articulated Inertia}
		\vspace{1mm}
		\State{$\hat{B}_k= \bar{B}_w$
		}\Comment{Knuckle 2 Articulated Bias}
		\vspace{1mm}
		\State{$\hat{M}_h= \Ad_{g_{hk}}^*\hat{M}_k\Ad_{g_{hk}}^{-1}$
		}\Comment{Knuckle 1 Articulated Inertia}
		\vspace{1mm}
		\State{$\hat{B}_h= \Ad_{g_{hk}}^*\bigl(\hat{M}_k(J_{hk}\ddot{q}_{hk}-\ad_{V_{hk}}\Ad_{g_{kh}}V_k)+\hat{B}_k\bigr)$
		}\Comment{Knuckle 1 Articulated Bias}
		\vspace{1mm}
		\State{$\bar{M}_h=\hat{M}_h-\dfrac{\hat{M}_hJ_{bh,z}J_{bh,z}^T\hat{M}_h}{J_{bh,z}^T\hat{M}_hJ_{bh,z}}$}
		\vspace{1mm}
		\State{$\bar{B}_h=\hat{B}_h-\dfrac{\hat{M}_hJ_{bh,z}(J_{bh,z}^T\hat{B}_h-\tau)}{J_{bh,z}^T\hat{M}_hJ_{bh,z}}-\bar{M}_h(\ad_{V_{bh}}V_h-\dot{J}_{bh,z}\dot{z}-\dot{J}_{bh,\delta}\dot{\delta}-J_{bh,\delta}\ddot{\delta})$}
		\vspace{1mm}
		\EndFor
		\vspace{1mm}
		\State{$\hat{M}_b=M_b+\displaystyle\sum_h\Ad_{g_{bh}}^*\bar{M}_h\Ad_{g_{bh}}^{-1}$}\Comment{Chassis Articulated Inertia}
		\vspace{1mm}
		\State{$\hat{B}_b=-W_{be}+\ad_{V_b}^*M_bV_b+\displaystyle\sum_h\Ad_{g_{bh}}^*\bar{B}_h$}\Comment{Chassis Articulated Bias}
		\vspace{1mm}
	\end{algorithmic}
\end{algorithm}

%Both inertia and bias are expressed in the body-fixed components.
In the case the DCC architecture is considered, the camber angle acceleration $\ddot{q}_{hk}$ is known, since $q_{hk(t)}$ is the assumed input, while the torque $\tau_{hk}$ has to be computed together with the rest of unknown accelerations, which are the outcome of the third sweep. This leads to a \emph{Hybrid Dynamics} problem that requires to carefully revisit the ABA's second and third steps.
The approach used is inspired by~\cite[Section~2.3]{Kim:notes:2012} where the the notation and conventions here employed were adapted according to~\cite{Domenighini:Designs:2023}.
%Specifically, apart from the labels assigned to the quantities, the primary difference between the two methods is the ordering of the twist vector. In this work, the twist vector is defined as $V = [v,\omega]^T$, whereas in the other method, it is defined as $V=[\omega,v]^T$. This difference might seem minor but leads to significant changes in the definitions of Adjoint operators, inertia matrices, bias and wrench vectors.

%For a generic articulated body the Newton-Euler equations can be expressed in the form:
%\begin{equation}\label{eqn:NE}
%	W=\hat{M}\dot{V}+\hat{B}.
%\end{equation}
%For a leaf node of the kinematic tree, having no children, the articulated inertia $\hat{M}$ is simply the generalized inertia $M$ of the body itself. Consistently, the bias $\hat{B}=B$, where:
%\begin{equation}
%	B=-W_e+\ad_V^*MV,
%\end{equation}
%and where $W_e$ is the resultant external wrench applied to the body and $\ad_V^*MV$ accounts for generalized gyroscopic and centrifugal forces, which are bi-linear in $V$. According to~\cite[Section~4.3.3]{Murray:book:1994} $\ad^*=-\ad^T$ has been defined. (If $v$ and $\omega$ are the translational and rotational component of $V$, then $\ad^*_V=\left[\begin{smallmatrix}\hat{\omega}&0\\\hat{v}&\hat{\omega}\end{smallmatrix}\right]$).

Starting from the leaf nodes, which in our case are represented by the wheels ($\sref{W}$), after providing information such as external wrenches, twists, joint geometry, body masses, joint forces / torques for traditional connections, and time evolution of variables defining kinematically imposed joints, the backward propagation of inertia and bias can be performed.

Finally the last step, denoted as \textbf{Step 3}, starting from the root node, computes joint variable accelerations and consequently body twists derivatives. For the kinematically controlled camber joints, the torques $\tau_{hk}$ needed to maintain the desired state of motion (DCC architecture) are computed as a by-product of the local inverse dynamics. This information is crucial to prevent actuators from  saturating or operating outside their bandwidth in the planned trajectories.

The root twist acceleration is determined by solving equation $W=\hat{M}\dot{V}+\hat{B}$, considering that the chassis is free-floating. The other unknown quantities, such as variable accelerations and torques, are computed projecting the Newton-Euler equations of each body along the respective Jacobian and then solving for the variable. See also~\cite{Domenighini:Designs:2023} for more details in the derivation of the equations.

\begin{algorithm}[h]
	\floatname{algorithm}{Step}
	\caption{Forward Propagation of Acceleration}\label{alg:step3}
	\begin{algorithmic}[1]
		\State{$\dot{V}_b=-\hat{M}_b^{-1}\hat{B}_b$}\Comment{Chassis Rigid-Body Acceleration}
		\vspace{1mm}
		\For{$h=h_1,h_2,h_3,h_4$}
		\vspace{1mm}
		\State{$\ddot{z}=-\dfrac{J_{bh,z}^T\bigl(\hat{B}_h+\hat{M}_h(\Ad_{g_{bh}}^{-1}\dot{V}_b-\ad_{V_{bh}}V_h+\dot{J}_{bh,z}\dot{z}+\dot{J}_{bh,\delta}\dot{\delta}+J_{bh,\delta}\ddot{\delta})\bigr)-\tau}{J_{bh,z}^T\hat{M}_hJ_{bh,z}}$}
		\vspace{1mm}
		\State{$\dot{V}_{bh}=\dot{J}_{bh,z}\dot{z}+\dot{J}_{bh,\delta}\dot{\delta}+J_{bh,z}\ddot{z}+J_{bh,\delta}\ddot{\delta}$}
		\vspace{1mm}
		\State{$\dot{V}_h=\Ad_{g_{bh}}^{-1}\dot{V}_b-\ad_{V_{bh}}V_h+\dot{V}_{bh}$}\Comment{Knuckle 1 Rigid-Body Acceleration}
		\vspace{1mm}
		\State{$\tau_{hk} = J_{hk}^T\bigl(\hat{B}_k+\hat{M}_k(\Ad_{g_{kh}}\dot{V}_h-\ad_{V_{hk}}V_k + J_{hk}\ddot{q}_{hk})\bigr)$}
		\vspace{1mm}
		\State{$\dot{V}_{bh} = J_{hk}\ddot{q}_{hk}$}
		\vspace{1mm}
		\State{$\dot{V}_k=\Ad_{g_{hk}}^{-1}\dot{V}_h-\ad_{V_{hk}}V_k+\dot{V}_{bh}$}\Comment{Knuckle 2 Rigid-Body Acceleration}
		\State{$\dot{\omega}=-\dfrac{J_{kw}^T\bigl(\hat{B}_w+\hat{M}_w(\dot{V}_k-\ad_{V_{kw}}V_w)\bigr)}{J_{kw}^T\hat{M}_wJ_{kw}}$}
		\vspace{1mm}
		\EndFor
		\vspace{1mm}
	\end{algorithmic}
\end{algorithm}

\subsection{Dynamic Model}
The dynamics equations, assembled via the procedure described  in Sec.~\ref{sec:aba}, are implemented in the software code by using a state-space representation $\dot{x}=f(x,u)$.
%\begin{equation}\label{eqn:f}
%	\dot{x}=f(x,u).
%\end{equation}
The state vector is $x = (q_{gb};V_{gb};z;\dot{z};\omega;q_{hk};\dot{q}_{hk})\in\bbR^{32}$, with components defined in previous sections. As already mentioned, in the model the camber joint is kinematically controlled as per the DCC architecture. Hence, the associated control vector is $u=(T_a;T_b;\delta;\ddot{q}_{hk})\in\bbR^{16}$, where each element is in $\bbR^{4}$ and collects the controls of the branch to each wheel. The torque is split into acceleration and braking components, allowing for minor modifications to model on-board motors coupled with traditional brakes and/or in-wheel motors as well. These components have their reaction torques in two different bodies in case of on-board motors and traditional brakes: the chassis and the knuckles, respectively. The $\delta$ vector collects the four translations of the steering rod attachment points, as described in Section~\ref{sec:suspanal}, while for the vector $\ddot{q}_{hk}$ no further explanation should be necessary. Clearly, this representation is valid for the model with active camber. The simpler model (no DCC) shares with the former the first 24 components of the state vector and the first 12 of the control vector.

The time derivative of the state vector has also been expressed previously except for the first element, $q_{gb}$. Its derivative is a function of the state and can be recovered using a standard procedure~\cite{Domenighini:Designs:2023} as follows:
\begin{equation}\label{eqn:dotqgb}
	\dot{q}_{gb}=J_{gb}^{-1}(q_{gb})V_{gb},
\end{equation}
where $J_{gb}(q_{gb})\in\bbR^{6\times6}$ is the Jacobian of the virtual joint connecting the ground to the chassis, and it is computed via the PoE parameterization of the transformation $g_{gb}$.
%, similarly to $J_{bh}$ in Eq.~\eqref{eqn:Jbh}.