### 2025-04-28 Fig/TikZ_Sources/tikz_covariance_evolution_matteo.tex

\begin{tikzpicture}[%
	smallnode/.style={%
		circle, draw, minimum size=3mm, inner sep=0pt
	},
	r_smallnode/.style={%
		circle, draw, minimum size=3mm, inner sep=0pt, red
	},
	g_smallnode/.style={%
		circle, draw, minimum size=3mm, inner sep=0pt, mygreen
	},
	every path/.style={->, thick},
	x=1.5cm, y=1.5cm
	]

	\newcommand{\Plab}[2]{\bP^{#2}_{#1}}
	\definecolor{mygreen}{RGB}{0 150 0}

	% Nodes
	% First row
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-2}\)}] (Pk-2H-2) at (1,0) {};
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-1}\)}] (Pk-1H-1) at (2,0) {};
	\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H}\)}] (PkH) at (3,0) {};
	\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{0}\)}] (Pk+10) at (4,0) {};
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{1}\)}] (Pk+21) at (5,0) {};

	\draw[bend left=20] (Pk-2H-2) to (Pk-1H-1);
	\draw[bend left=20] (Pk-1H-1) to (PkH);
	\draw[bend left=20] (Pk+10) to (Pk+21);

	\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k-H}{0}\)}] (Pk-H0) at (0,0) {};
	\node[label={center:\(\dots\)}] (dots10) at (-.5, 0) {};
	\node[label={center:\(\dots\)}] (dots11) at (.5, 0) {};
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+H}{H-1}\)}] (Pk+HH-1) at (6,0) {};
	\node[label={center:\(\dots\)}] (dots12) at (5.5, 0) {};
	\node[label={center:\(\dots\)}] (dots13) at (6.5, 0) {};
	% Second row
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-3}\)}] (Pk-2H-3) at (1,1) {};
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-2}\)}] (Pk-1H-2) at (2,1) {};
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H-1}\)}] (PkH-1) at (3,1) {};
	\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{H}\)}] (Pk+1H) at (4,1) {};
	\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{0}\)}] (Pk+20) at (5,1) {};

	\draw[bend left=20] (Pk-2H-3) to (Pk-1H-2);
	\draw[bend left=20] (Pk-1H-2) to (PkH-1);
	\draw[bend left=20] (PkH-1) to (Pk+1H);

	\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-H}{H}\)}] (Pk-HH) at (0,1) {};
	\node[label={center:\(\dots\)}] (dots20) at (-.5, 1) {};
	\node[label={center:\(\dots\)}] (dots21) at (.5, 1) {};
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+H}{H-2}\)}] (Pk+HH-2) at (6,1) {};
	\node[label={center:\(\dots\)}] (dots22) at (5.5, 1) {};
	\node[label={center:\(\dots\)}] (dots23) at (6.5, 1) {};

	% dots row
	\node[label={center,rotate=90:\(\dots\)}] (dots1) at (0, 1.7) {};
	\node[label={center,rotate=90:\(\dots\)}] (dots2) at (1, 1.7) {};
	\node[label={center,rotate=90:\(\dots\)}] (dots3) at (2, 1.7) {};
	\node[label={center,rotate=90:\(\dots\)}] (dots4) at (3, 1.7) {};
	\node[label={center,rotate=90:\(\dots\)}] (dots5) at (4, 1.7) {};
	\node[label={center,rotate=90:\(\dots\)}] (dots6) at (5, 1.7) {};
	\node[label={center,rotate=90:\(\dots\)}] (dots7) at (6, 1.7) {};

	% Third row
	\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H}\)}] (Pk-2H) at (1,2) {};
	\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{0}\)}] (Pk-10) at (2,2) {};
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{1}\)}] (Pk1) at (3,2) {};
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{2}\)}] (Pk+12) at (4,2) {};
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{3}\)}] (Pk+23) at (5,2) {};

	\draw[bend left=20] (Pk-10) to (Pk1);
	\draw[bend left=20] (Pk1) to (Pk+12);
	\draw[bend left=20] (Pk+12) to (Pk+23);

	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-H}{2}\)}] (Pk-H2) at (0,2) {};
	\node[label={center:\(\dots\)}] (dots30) at (-.5, 2) {};
	\node[label={center:\(\dots\)}] (dots31) at (.5, 2) {};
	\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+H}{0}\)}] (Pk+H0) at (6,2) {};
	\node[label={center:\(\dots\)}] (dots32) at (5.5, 2) {};
	\node[label={center:\(\dots\)}] (dots33) at (6.5, 2) {};

	% Fourth row
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-1}\)}] (Pk-2H-1) at (1,3) {};
	\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H}\)}] (Pk-1H) at (2,3) {};
	\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{0}\)}] (Pk0) at (3,3) {};
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{1}\)}] (Pk+11) at (4,3) {};
	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{2}\)}] (Pk+22) at (5,3) {};

	\draw[bend left=20] (Pk-2H-1) to (Pk-1H);
	\draw[bend left=20] (Pk0) to (Pk+11);
	\draw[bend left=20] (Pk+11) to (Pk+22);

	\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-H}{1}\)}] (Pk-H1) at (0,3) {};
	\node[label={center:\(\dots\)}] (dots40) at (-.5, 3) {};
	\node[label={center:\(\dots\)}] (dots41) at (.5, 3) {};
	\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k+H}{H}\)}] (Pk+HH) at (6,3) {};
	\node[label={center:\(\dots\)}] (dots42) at (5.5, 3) {};
	\node[label={center:\(\dots\)}] (dots43) at (6.5, 3) {};

	% Dashed rectangle
	\draw[dashed, thick, rounded corners] (2.6,-.25) rectangle (3.4,3.5);

\end{tikzpicture}