### 2025-06-04 Images/test0.tex
[mg] general reviews

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
	%uncomment if require: \path (0,470); %set diagram left start at 0, and has height of 470

	%Curve Lines [id:da3902337050237109] 
	\draw    (131,68.6) .. controls (115,164.6) and (172,202.6) .. (201,210.6) .. controls (230,218.6) and (301.88,256.97) .. (321,278.6) .. controls (340.13,300.22) and (365,360.6) .. (467,351.6) ;
	%Curve Lines [id:da9978493675513382] 
	\draw    (224,42.6) .. controls (208,138.6) and (250,136.6) .. (279,144.6) .. controls (308,152.6) and (368.88,190.97) .. (388,212.6) .. controls (407.13,234.22) and (417,285.6) .. (519,276.6) ;
	%Curve Lines [id:da9577580542618884] 
	\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ] [dash pattern={on 4.5pt off 4.5pt}]  (205,56.6) .. controls (189,152.6) and (231,150.6) .. (260,158.6) .. controls (289,166.6) and (349.88,204.97) .. (369,226.6) .. controls (388.13,248.22) and (405,304.6) .. (507,295.6) ;
	%Curve Lines [id:da8960204057983273] 
	\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ] [dash pattern={on 4.5pt off 4.5pt}]  (155,69.6) .. controls (139,165.6) and (191,186.6) .. (220,194.6) .. controls (249,202.6) and (324.88,251.97) .. (344,273.6) .. controls (363.13,295.22) and (362,332.6) .. (469,338.6) ;
	%Curve Lines [id:da018938305669631506] 
	\draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ] [dash pattern={on 0.84pt off 2.51pt}]  (176,61.6) .. controls (160,157.6) and (205,165.6) .. (234,173.6) .. controls (263,181.6) and (325.88,218.97) .. (345,240.6) .. controls (364.13,262.22) and (383,331.6) .. (485,322.6) ;
	%Straight Lines [id:da3684832566848931] 
	\draw    (288.9,137.15) -- (274.98,159.6) ;
	\draw [shift={(273.4,162.15)}, rotate = 301.8] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
	%Straight Lines [id:da19558075317267187] 
	\draw    (245,102.6) -- (233.97,120.59) ;
	\draw [shift={(232.4,123.15)}, rotate = 301.51] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;

	% Text Node
	\draw (229,83.4) node [anchor=north west][inner sep=0.75pt]    {$\beta =0$};
	% Text Node
	\draw (281.5,114.4) node [anchor=north west][inner sep=0.75pt]    {$\beta =0.1$};

\end{tikzpicture}