### 2025-06-04 Images/test1.tex

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,scale=0.5]
	%uncomment if require: \path (0,470); %set diagram left start at 0, and has height of 470

	%Curve Lines [id:da3902337050237109] 
	\draw    (209.6,68.4) .. controls (145.12,155.42) and (212.32,229.5) .. (271.54,259.19) .. controls (275.38,261.12) and (279.19,262.86) .. (282.93,264.4) .. controls (344.27,289.73) and (346.27,313.07) .. (354.93,333.73) ;
	%Straight Lines [id:da3684832566848931] 
	\draw    (336.93,189.73) -- (316.45,213.94) ;
	\draw [shift={(314.52,216.23)}, rotate = 310.23] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
	%Straight Lines [id:da19558075317267187] 
	\draw    (299.6,121.07) -- (263.8,137.25) ;
	\draw [shift={(261.07,138.48)}, rotate = 335.68] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
	%Straight Lines [id:da7674435216299973] 
	\draw    (354.93,333.73) -- (432.27,319.07) ;
	%Curve Lines [id:da6299814455818392] 
	\draw    (277.6,91.07) .. controls (238.93,153.73) and (274.93,179.73) .. (322.27,203.73) .. controls (369.6,227.73) and (424.27,267.07) .. (432.27,319.07) ;
	%Straight Lines [id:da357006401572867] 
	\draw    (209.6,68.4) -- (277.6,91.07) ;
	%Curve Lines [id:da08014732331041308] 
	\draw  [dash pattern={on 4.5pt off 4.5pt}]  (264.27,87.07) .. controls (225.6,149.73) and (260.27,187.07) .. (309.6,212.4) .. controls (358.93,237.73) and (410.27,269.73) .. (418.27,321.73) ;
	%Curve Lines [id:da8971212017686895] 
	\draw  [dash pattern={on 4.5pt off 4.5pt}]  (221.6,73.07) .. controls (151.6,161.73) and (245.6,240.4) .. (296.27,258.4) .. controls (346.93,276.4) and (357.6,305.73) .. (369.6,331.73) ;
	%Curve Lines [id:da47999458459142474] 
	\draw  [dash pattern={on 3.75pt off 3pt on 1.5pt off 3pt}]  (243.6,79.73) .. controls (182.93,173.07) and (272.93,223.73) .. (321.6,246.4) .. controls (370.27,269.07) and (384.27,297.07) .. (393.6,326.4) ;

	% Text Node
	\draw (285.67,100.73) node [anchor=north west][inner sep=0.75pt]    {$\beta =0$};
	% Text Node
	\draw (321.5,168.4) node [anchor=north west][inner sep=0.75pt]    {$\beta =0.1$};

\end{tikzpicture}